\documentclass[justified]{tufte-handout} 
\usepackage{amsfonts, amssymb, stmaryrd, fitch, natbib, qtree}
\usepackage{linguex, color, setspace, graphicx}
\usepackage{enumitem}
\usepackage{bussproofs}
\usepackage{turnstile}
\usepackage[super]{nth}
\thispagestyle{plain}
\definecolor{darkred}{rgb}{0.7,0,0.2}
\bibpunct{(}{)}{,}{a}{}{,}

\input xy
 \xyoption{all}

%New Symbols
\DeclareSymbolFont{symbolsC}{U}{txsyc}{m}{n}
\DeclareMathSymbol{\strictif}{\mathrel}{symbolsC}{74}
\DeclareMathSymbol{\boxright}{\mathrel}{symbolsC}{128}
\DeclareMathSymbol{\Diamondright}{\mathrel}{symbolsC}{132}
\DeclareMathSymbol{\Diamonddotright}{\mathrel}{symbolsC}{134}
\DeclareMathSymbol{\Diamonddot}{\mathord}{symbolsC}{144}
\renewcommand{\labelitemi}{$\triangleright$}
\renewcommand{\labelitemii}{$\circ$}
\renewcommand{\labelitemiii}{$\triangleright$}

%New commands
\newcommand{\bitem}{\begin{itemize}}
\newcommand{\eitem}{\end{itemize}}
\newcommand{\lang}{$\langle$}
\newcommand{\rang}{$\rangle$}
\newcommand{\back}{$\setminus$}
\newcommand{\HRule}{\rule{\linewidth}{0.1mm}}
\newcommand{\llm}[2][]{$\llbracket${#2}$\rrbracket^{#1}$}
\newcommand{\ul}{$\ulcorner$}
\newcommand{\ur}{$\urcorner\ $}
\newcommand{\urn}{$\urcorner$}
\newcommand{\sub}[1]{\textsubscript{#1}}
\newcommand{\sups}[1]{\textsuperscript{#1}}
\newtheorem{proposition}{\textbfb{Proposition}}[section]
\newtheorem{definition}[proposition]{\textbf{Definition}}
\newcommand{\bfw}{\begin{fullwidth}}
\newcommand{\efw}{\end{fullwidth}}

\begin{document}

\begin{fullwidth}
\noindent\LARGE Newcomb's Problem \& Probability \normalsize \\[.3cm]
\noindent  \textsc{24.118 Recitation Section $\bullet$ Matthias Jenny\\  {\texttt{\href{mailto:mjenny@mit.edu}{mjenny@mit.edu}}} $\bullet$ Office:  32-D927 $\bullet$ Hours: Thu 11:30-12:30} \hfill{September 26, 2014}
\noindent\HRule
\end{fullwidth}

\section{Three simple cases}

\begin{enumerate}
\item Suppose you get to class and there's a pop quiz with one simple `yes' or `no' question. You haven't looked at this week's lecture material. What should you do? Try to answer using the concepts of Expected Value Maximization and the Principle of Indifference.\end{enumerate}

\noindent Notes:  \underline{\hspace{15.75cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\

\begin{enumerate}
\setcounter{enumi}{1}
\item Suppose you're in a Newcomb situation with an 80\% percent accurate predictor. How would you defend the option of TwoBoxing.
\end{enumerate}

\noindent Notes:  \underline{\hspace{15.75cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\

\begin{enumerate}
\setcounter{enumi}{2}
\item Suppose you're in a repeated Newcomb situation and every time you decide whether to OneBox or to TwoBox you flip a coin to make your decision. Is it possible for the predictor to be 80\% accurate?
\end{enumerate}

\noindent Notes:  \underline{\hspace{15.75cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\

\section{Pset 3, Problem 1}

\begin{quote}
Suppose that we live an a deterministic world, and that I am about to toss an ordinary coin. What is the objective probability that the toss will land heads?
\end{quote}

\noindent\emph{Reminder:} The Objective-Subjective Connection says that the objective probability of an event at time $t$ is the subjective probability that a perfectly rational agent would assign to that event, if she had access to all evidence which is admissible at $t$.\\

\noindent Notes:  \underline{\hspace{15.75cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\

\section{Pset 3, Problems 3 \& 4}

\begin{quote}
You are offered a chance to play a game that works as follows. A coin will be tossed until it lands heads. (If it never lands heads, the game is off.) If the coin lands heads on the $n^\mathrm{th}$ toss, you receive $\frac{2^n}{4}$.

\begin{itemize}\item What is the expected value of playing the game?
\item How much should one be willing to pay for the privilege of playing the game?\end{itemize}
\end{quote}

\noindent Notes:  \underline{\hspace{15.75cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\\\\underline{\hspace{16.88cm}}\\


\end{document}
