%issue with tkz-berge.sty and tkz-graph being removed from 2020 release of texlive:
% https://tex.stackexchange.com/questions/547173/texlive-2020-how-to-deal-with-tikz-berge-and-tikz-kiviat-packages-that-were-rem
% just downloaded the .sty files and added them to local folder i'm compiling this doc in! 

%https://www.youtube.com/watch?v=Lli99OmkPwM

%i guess if they get to drop a Pset, I don't want each to count as 9 points...want a number divisible by 8 lol; or I could do each is worth 10 points and give them a freebie? or play with the participation some more! 

\documentclass[11pt, a4paper]{article}
\usepackage[inner=2cm,outer=2cm,top=2.3cm,bottom=1.5cm]{geometry}
\pagestyle{empty}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{bm}
\usepackage{multicol}
\usepackage{amssymb}
\usepackage{amsmath}

%\usetikzlibrary{snakes}
%\usetikzlibrary{lindenmayersystems}
\usetikzlibrary{arrows,petri,topaths}
\usetikzlibrary{plotmarks}
\usepackage{tkz-berge}
\usepackage[position=top]{subfig}
\usepackage{enumitem}
\usepackage{fancyhdr, lastpage, bbding, pmboxdraw}
%\usepackage[usenames,dvipsnames]{color}

\definecolor{darkblue}{rgb}{0,0,.6}
\definecolor{darkred}{rgb}{.7,0,0}
\definecolor{darkgreen}{rgb}{0,.6,0}
\definecolor{red}{rgb}{.98,0,0}
\usepackage[colorlinks,pagebackref,pdfusetitle,urlcolor=darkblue,citecolor=darkblue,linkcolor=darkred,bookmarksnumbered,plainpages=false]{hyperref}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\pagestyle{fancyplain}
\fancyhf{}
%\lhead{ \fancyplain{}{24.118 \ Paradox and Infinity} }
\lhead{ \fancyplain{}{24.118 \ Health and Wellness} }
%\chead{ \fancyplain{}{} }
\rhead{ \fancyplain{}{Spring 2023} }
%\rfoot{\fancyplain{}{page \thepage\ of \pageref{LastPage}}}
\fancyfoot[RO, LE] {page \thepage\ of \pageref{LastPage} }
\thispagestyle{plain}

%%%%%%%%%%%% LISTING %%%
\usepackage{listings}
\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}
\usepackage{verbatim} % used to display code
\usepackage{fancyvrb}
\usepackage{acronym}
\usepackage{amsthm}
\VerbatimFootnotes % Required, otherwise verbatim does not work in footnotes!



\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{CadetBlue}{cmyk}{0.62,0.57,0.23,0}
\definecolor{lightlightgray}{gray}{0.93}



\lstset{
%language=bash,                          % Code langugage
basicstyle=\ttfamily,                   % Code font, Examples: \footnotesize, \ttfamily
keywordstyle=\color{OliveGreen},        % Keywords font ('*' = uppercase)
commentstyle=\color{gray},              % Comments font
numbers=left,                           % Line nums position
numberstyle=\tiny,                      % Line-numbers fonts
stepnumber=1,                           % Step between two line-numbers
numbersep=5pt,                          % How far are line-numbers from code
backgroundcolor=\color{lightlightgray}, % Choose background color
frame=none,                             % A frame around the code
tabsize=2,                              % Default tab size
captionpos=t,                           % Caption-position = bottom
breaklines=true,                        % Automatic line breaking?
breakatwhitespace=false,                % Automatic breaks only at whitespace?
showspaces=false,                       % Dont make spaces visible
showtabs=false,                         % Dont make tabls visible
columns=flexible,                       % Column format
morekeywords={__global__, __device__},  % CUDA specific keywords
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{center}
{\Large Health and Wellness Resources}\\
{  \textsc{24.118 \ Paradox and Infinity}}
\end{center}
%\date{September 26, 2014}

As my dad always says, ``Your health is your greatest wealth!'' And that includes your mind! Almost everyone goes through a lot of stuff in their lives that can be difficult to handle. I encourage you to think freely and talk openly about the adversities you are going through and try to take appropriate steps to manage them on a regular basis. I am not a mental health professional, but I am a sympathetic ear. I encourage you to take advantage of the bevy of resources available to you, some of which are listed below. 
%(rather than letting things boil over or go unchecked)

If your adversities ever relate to school, try to remember that school is just school and that you are the arbiter of your own success. If your adversities relate at all to relationships, try to remember that there are many places and people in this great big sea of humanity to look for support. In any case, many people take time off and go on to live superbly happy, fulfilling lives! You've got a multitude of options, many more than you realize now! Try not to ever feel stuck in a corner! 
%If your adversities relate at all to relationships, try to remember that there are A TON of people in this great big beautiful sea of humanity
%If your adversities relate at all to relationships, try to remember that there are a ton of people in this great big sea of humanity 
%If your adversities relate at all to relationships, try to remember that there are many places to look for support.

More than anything, I hope you can remember the following facts: \textit{You are enough. You are loved.} 
%in times of crisis

\vskip.15in

\noindent\textbf{Links and deets:}

\begin{itemize}
    \item \href{https://medical.mit.edu/services/mental-health-counseling/at-a-glance}{\textit{Student Mental Health \& Counseling Services}}
    
    \begin{itemize}
        \item Phone number: 617-253-2916. 
        \item ``Mental health clinicians are available 24 hours a day for urgent matters.''
        
        \item Provides evaluations, consultations, treatment, referrals!
    \end{itemize}
    
    \item \href{https://medical.mit.edu/services/mental-health-counseling/self-care-resources}{\textit{Self-care Resources}!}
    
    \begin{itemize}
        \item Lots of helpful videos and podcasts addressing common stressors! \\ (e.g. imposter syndrome, anxiety, procrastination, perfectionism, etc.)
    \end{itemize}
    
    \item \href{https://studentlife.mit.edu/aods#}{\textit{Alcohol and Other Drug Services (AODS)}}
    
    \begin{itemize}
        \item Support for issues related to alcohol and other drug abuse, stress, and violence.
        \item Substance addiction can be incredibly difficult. I would take this seriously!
        %\item Substance addiction is no joke! Take back your life! 
    \end{itemize}
    
    \item \href{https://studentlife.mit.edu/vpr}{\textit{VPR (Violence Prevention and Response)}}
    
    \begin{itemize}
        \item 617-253-2300 (9am--5pm Monday thru Friday or afterhours to be directed to 24/7 resources)
       \item Support for individuals who are dealing with issues such as sexual assault, dating and domestic violence, stalking, and sexual harassment. 
    \item ``If you are unable to speak safely in your current situation, scheduled VPR advocacy appointments can be text- or chat-based.''
     
        \item Hotline available 24 hours a day during the school year
       % \item These are absolutely devastating and gut-wrenching issues. \textit{Always} show your fellow humans the respect that they deserve.
        \item This should go without saying, but \textit{please please please} do not be part of the problem on these absolutely devastating and gut-wrenching issues. \textit{Always} show your fellow humans the respect that they deserve.
        \item Additionally, if you see someone who IS part of the problem and feel safe doing so, speak up! We must never tolerate this egregious behavior from \textit{anyone}. 
    \end{itemize}

    
    \item \href{https://medical.mit.edu/my-mit/parents/mental-health-resources}{\textit{A list of some additional resources}!}
    
    \item If you've had experience with additional resources that you think belong here, please let me know so that I can add them!
\end{itemize}

%As one of these websites wisely says, 

\newpage


\noindent\textbf{Motivational(?) Quotes!}

\begin{quote}
    ``One will be imprisoned in a room with a door that is unlocked and opens inwards; \\ \phantom{vvvv} as long as it does not occur to them to pull rather than push it.'' \\
   \phantom{vvvvvv} --Ludwig Wittgenstein (in \textit{Culture and Value})
   
   \vskip.15in
    
     ``It is a sign of strength to recognize these kinds of concerns and make a plan \\ \phantom{vvvv} to deal with them constructively.'' \\ \phantom{vvvvvv} -- MIT Mental Health \& Counseling
     
       \vskip.15in
     
     ``We often don’t realize we’ve gone downhill until we have to go up again. \\ \phantom{vvvv} We’ll all get there in our own way.''  --Your instructor 
     
       \vskip.15in
    
     ``The future belongs to those who believe in the beauty of their dreams.'' \\ \phantom{vvvvvv} --Eleanor Roosevelt
     
       \vskip.15in
     
     ``If you're not failing at something, you're not trying hard enough. Failing is part of it."  \\ \phantom{vvvvvv} --Johnny Knoxville 
     
       \vskip.15in
     
     ``If everything is under control, you’re not going fast enough.”  \\ \phantom{vvvvvv} --Mario Andretti
     
       \vskip.15in
     
     ``Specialization is for insects.” (but hey, insects are crushing it!) \\ \phantom{vvvvvv} --Robert Heinlein
     %last two are quoted by Landsmann in his email signature 
     
    
\end{quote}


\end{document}