
\documentclass[12pt,a4paper]{article}
\usepackage{pset-2018}
%\usepackage{wasysym}
%\usepackage{marvosym}
\usepackage{mathabx}



%Questions and Answers
\qa{a} % q ="questions only"; a="answers only";  b="both"
\usepackage{qa}


\begin{document}

\psintro{Problem Set 1: Infinite Cardinalities}



% Idea for next time: ask them to show that Cantor's Theorem shows that the powerset axiom is incompatible with the existence of a universal set.



\subsection*{Part I} 
\begin{enumerate}


\com{
\item 
\question{Let \(A\) be a set and let \(R\) be a relation that holds amongst members of that set. Then: 
  \begin{itemize}
\item   
   \(R\) is \textbf{reflexive} on \(A\) if and only if: for any \(a\in A\), \(aRa\). 
      \item  \(R\) is \textbf{symmetric} on \(A\) if and only if: for any \(a,b\in A\), if \(aRb\) then \(bRa\). 
  \item  \(R\) is \textbf{transitive} on \(A\) if and only if: for any \(a,b,c\in A\), if \(aRb\) and \(bRc\) then \(aRc\). 
\end{itemize}
(\emph{Example:} Let \(A\) be the set of people, and let \(R\) be the relation ``having the same birthday as". \(R\) is reflexive, since everyone has the same birthday as herself. \(R\) is symmetric, since whenever \(a\) has the same birthday as \(b\), \(b\) will have the same birthday as \(a\). And \(R\) is transitive: for any three people \(a\), \(b\), and \(c\), if \(a\) has the same birthday as \(b\), and \(b\) has the same birthday as \(c\), then \(a\) has the same birthday as \(c\).)}

\begin{enumerate}
\item  \question{Consider the relation ``is less than or equal to". Is this relation: $(i)$ reflexive, $(ii)$ symmetric, and $(iii)$ transitive on the set of natural numbers? (5 points.)}

\answer{ ``Less than or equal to" is reflexive on the set of natural numbers, since every number is less than or equal to itself. It is also transitive on the natural numbers, since if \(x\) is less than or equal to \(y\), and \(y\) is less than or equal to \(z\), \(x\) is less than or equal to \(z\). It is not, however, symmetric on the natural numbers, since if \(x\) is less than or equal to \(y\), there is no guarantee that \(y\) is less than or equal to \(x\). That is true only in the special case where \(x\) is equal to \(y\).}
  
 \item    \question{There are ten guests at a dinner party, sitting around a large table. Consider the relation \(R\) such that guest \(a\) bears \(R\) to guest \(b\) if and only if \(a\)'s seat is immediately adjacent to \(b\)'s seat. Is \(R\): $(i)$ reflexive, $(ii)$ symmetric, and $(iii)$ transitive on the set of guests?  (Keep in mind that no seat is adjacent to itself.) (5 points.)}
 
 \answer{
 \(R\) fails to be reflexive because nobody is in a seat immediately adjacent to her own seat (your seat is only immediately adjacent to your neighbor's seats).

\(R\) also fails to be transitive; to see this, suppose that \(a\) is seated immediately to the right of \(b\) and that \(b\) is seated immediately to the right of \(c\); then \(a\) bears \(R\) to \(b\) and \(b\) bears \(R\) to \(c\), but \(a\) doesn't bear \(R\) to \(c\).

On the other hand, \(R\) is symmetric, since the only way for \(a\) to be in a seat immediately adjacent to \(b\) is for \(b\) to be in a seat immediately adjacent to \(a\). 
 }


\end{enumerate}
}



\item \question{A \textbf{function} $f : A \rightarrow B$ is an assignment of each member of $A$ to some member of $B$. A function $f: A \rightarrow B$ is an \textbf{injection}  if $f(x) \neq f(y)$ whenever $x \neq y$  ($x,y \in A$). A function $f: A \rightarrow B$ is a \textbf{surjection} if there is no member of $B$ to which $f$ fails to assign some member of $A$. 

For each of the following functions $f: A \rightarrow B$ determine whether $f$ is injective and whether it is surjective.}

\begin{enumerate}

% Removed for variety
\com{
\item\question{
(3 points)
\[
\begin{array}{ccl}
f(x) &= &2x\\
A &= &\mathbb{N}\\
B &= &\mathbb{N}
\end{array}
\] }

\answer{The function is injective, since no two elements of \(A\) are mapped to the same element of $B$. It is not surjective, since not every element of \(B\) is hit by an element of $A$. For instance, there is no number in set $A$ whose square is 69 but 69 is an element of set \(B\).  }
}


\item\question{
(2 points)
\[
\begin{array}{ccl}
f(x) &= &x+1\\
A &= &\mathbb{N}\\
B &= &\mathbb{N}
\end{array}
\] }

\answer{The function is injective, since no two elements of \(A\) are mapped to the same element of $B$. It is not surjective, since zero is in \(B\) but is not hit by an element of $A$.}


% Left out for variety
\com{
\item\question{
(2 points)
\[
\begin{array}{ccl}
f(x) &= &x+1\\
A &= &\mathbb{N}\\
B &= &\mathbb{N}-\set{0}
\end{array}
\] }

\answer{The function is injective, since no two elements of \(A\) are mapped to the same element of $B$. It is also surjective, since every element of \(B\) is hit by an element of $A$.}

}


\item\question{
(2 points)
\[
\begin{array}{ccl}
f(x) &= &x+1\\
A &= &\mathbb{Z}\\
B &= &\mathbb{Z}
\end{array}
\] }

\answer{The function is injective, since no two elements of \(A\) are mapped to the same element of $B$. It is also surjective, since every element of \(B\) is hit by an element of $A$.}





\item\question{
(2 points)
\[
\begin{array}{ccl}
f(x) &= &x^2\\
A &= &\mathbb{Z}\\
B &= &\mathbb{Z}
\end{array}
\] }

\answer{The function is not injective, since, e.g.~$-1$ and $1$ are both mapped to $1$. It is not surjective, since $-1$ is in \(B\) but is not hit by an element of $A$.}





\item   \question{
(2 points)
\[
\begin{array}{ccl}
f(x) &= &x+\pi\\
A &= &\mathbb{R}\\
B &= &\mathbb{R}
\end{array}
\] }


\answer{ The function is injective and surjective. Every element of set \(A\) is paired with exactly one element of set \(B\), and every element of set \(B\) is paired with exactly one element of set \(A\). 0 is paired with 1, $-\pi$ is paired with $-\pi+1$, and so on.}






% Removed for variety
\com{
\item   \question{
(3 points)
\[
\begin{array}{ccl}
f(x) &= &x-2\\
A &= &\mathbb{Z}\\
B &= &\mathbb{Z}
\end{array}
\] }


\answer{ The function is injective and surjective. Every element of set \(A\) is paired with exactly one element of set \(B\), and every element of set \(B\) is paired with exactly one element of set \(A\).}
}


\end{enumerate}
\question{\emph{Notation:} $\mathbb{N} = \{0,1,2,\dots\}$ is the set of natural numbers, $\mathbb{Z} = \{\dots -2,-1,0,1,2,\dots\}$ is the set of integers, and $\mathbb{R}$ is the set of real numbers.}






\item \question{A \textbf{bijection} from set $A$ to set $B$ is a function from $A$ to $B$ which is both injective and surjective. Between which of the following pairs of infinite sets is it possible to construct a bijection?}

\begin{enumerate}

\item \question{The set of negative integers and the set of non-negative integers. (2~points.)}

\answer{Yes.}


\item \question{The set of prime numbers and the set of real numbers. (2~points.)}

\answer{No.}



\item \question{The set of prime numbers and the set of real numbers between 0 and 1. (2~points.)}

\answer{No.}

\item \question{The rational numbers and the set of rational numbers between 0 and 1. (2~points.)}

\answer{Yes.}

\end{enumerate}



\item \question{The following principles give conflicting answers to cardinality questions:


\begin{description}
\item[The Proper Subset Principle]
Suppose $A$ is a {proper subset} of $B$. Then $A$ and $B$ are \emph{not} of the same size: $B$ has more members than $A$.
\end{description}
\begin{description}
\item[The Bijection Principle]
Set $A$ has the same size as set $B$ if and only if there is a \emph{bijection} from $A$ to $B$.
\end{description}
For each of the questions below determine which of the following answers is correct: ``yes'', ``no'', or ``not determined''.
}




\begin{enumerate}

\item \question{Are $\set{0,1,2}$ and $\set{0,1,2,3}$ of the same size, according to the Proper Subset Principle? Are they of the same size according to the Bijection Principle? (3~points.)
}

\answer{No: $\set{0,1,2} \subsetneq \set{0,1,2,3}$

No: there is no bijection.}



\item \question{Are $\set{0,1,2}$ and $\set{1,2,3,4}$ of the same size, according to the Proper Subset Principle? Are they of the same size according to the Bijection Principle? (3~points.)
}

\answer{Answer not determined by the Proper Subset Principle, since neither set is a proper subset of the other.

No: there is no bijection.}



\item \question{Are the set of prime numbers and the set of natural numbers of the same size, according to the Proper Subset Principle? Are they of the same size according to the Bijection Principle? (3~points.)
}

\answer{No: $\set{n : n \text{ is prime}} \subsetneq \mathbb{Z}$

Yes: there is a bijection.}

\end{enumerate}



\question{\fbox{\parbox{150mm}{\emph{Reminder:} Although you'll need to think about the Proper Subset Principle for the purposes of this question, it won't be relevant for the rest of the problem set. In this class we follow Cantor---and current mathematical practice---in rejecting the Proper Subset Principle and assessing cardinality questions on the basis of the Bijection Principle.}}}








\item \question{
Every room in Hilbert's Hotel is occupied. New guests show up.
}

\begin{enumerate}

\item There is one new guest for each rational number. Can all of them be accommodated without asking anyone to share a room? (3~points.)

\answer{Yes. There are countably many rational numbers, and countably many spaces can be freed up in Hilbert's Hotel by asking guest $n$ to move to room $2n$.}

\item There is one new guest for each real number. Can all of them be accommodated without asking anyone to share a room? (3~points.)
 

\answer{No. There are only countably many hotel rooms. If we could fit uncountably many guests in the hotel, then there would be a bijection between a countable set (the set of hotel rooms) and an uncountable set (the set of guests). But there is no such bijection. So an uncountable number of guests will not fit.}


\end{enumerate}




\item \question{
Say that a sequence of objects is \textbf{dense} if and only if between any two members of the sequence there is a third. Must a sequence have more elements than there are natural numbers in order to be dense? Consider only sequences with at least two elements. (3~points.)
}

\answer{No. The sequence of rational numbers is dense, but there are as many rational numbers as natural numbers, since a bijection can be constructed between the rational numbers and the natural numbers.}




%Left out for variety
\com{

\item \question{Describe a set that contains no integers but has the same cardinality as the set of integers. (4~points.)}

\answer{There are many such sets. Here is one example: $\set{x+\pi : x \in \mathbb{N}}$.}
}

\end{enumerate}





%%%%%%%%%%%%%%%
%PART II
%%%%%%%%%%%%%%%%
\subsection*{Part II} 
\begin{enumerate}
  \setcounter{enumi}{5}
%%%%%%%%%%%%%%%

% Left out for variety
\com{
\item \question{Show that there is a bijection between the set of integers \{\ldots -2, -1, 0, 1, 2, \ldots\} and the set of squares of integers $\{0, 1, 4, 9, 16, \dots\}$. (5 points)}

\answer{$f(n) = n^2$ is a bijection from the natural numbers to their squares. But $g$ is a bijection from the natural numbers to the integers:

\begin{itemize}

\item[] $g(n) = n/2$ (if $n$ is even)

\item[] $g(n) = -(n+1)/2$ (if $n$ is odd)

\end{itemize}
So $f(g^{-1}(n))$ is a bijection from the integers to their squares.
}
}

\item \question{Show that there is a bijection from $\mathbb{Z}$ to the set of powers of seven $\set{7^1,7^2,7^3,\dots}$. (4~points)}

\answer{$f(n) = 7^{n+1}$ is a bijection from the natural numbers to $\set{7^1,7^2,7^3,\dots}$. And $g$ is a bijection from the natural numbers to the integers:

\begin{itemize}

\item[] $g(n) = n/2$ (if $n$ is even)

\item[] $g(n) = -(n+1)/2$ (if $n$ is odd)

\end{itemize}
So $f(g^{-1}(n))$ is a bijection from the integers to $\set{7^1,7^2,7^3,\dots}$.
}



\item  \question{The Cantor-Schroeder-Bernstein Theorem states that if there is an injection from $A$ to $B$ and an injection from $B$ to $A$, then there is a bijection from $A$ to $B$.}
\begin{enumerate}

\item \question{Show that there is an injection from $\mathbb{N}$ to the set of prime numbers  $\set{2,3,5,7,\dots}$. (5~points) 

(You may help yourself to the observation that whenever $p_1,\dots,p_n$ are primes, there is a smallest prime greater than each of $p_1,\dots,p_n$.) }

\answer{Let $f(0) = 2$ and, for any $n \in \mathbb{N}$, let $f(n+1)$ be the smallest primer greater than $f(0), \dots, f(n)$. We can use mathematical induction to verify that $f$ is well-defined for each $n \in \mathbb{N}$. (The base case is trivial; the inductive case follows from the observation that whenever $f(0),\dots,f(n)$ are primes, there is a smallest prime greater than $f(0),\dots,f(n)$.) 

To verify that $f$ is an injection from $\mathbb{N}$ to the set of prime numbers, we need to show that $f(n)$ is always a prime number and that $f(n) \neq f(m)$ whenever $n \neq m$. The former is trivial; the latter follows immediately from the observation that the definition of  $f$ guarantees that $f(n) < f(m)$ whenever $n < m$.


}

\item \question{Use the Cantor-Schroeder-Bernstein Theorem to show that there is a bijection from $\mathbb{N}$ to the set of prime numbers. (5~points)}

\answer{
The identity function is an injection from the prime numbers to the natural numbers. And it follows from part $(a)$ that there is an injection from the natural numbers to the prime numbers. So, by the Cantor-Schroeder-Bernstein Theorem, there is a bijection from the natural numbers to the prime numbers.
}

\end{enumerate}



%Left out for variety
\com{

\item \question{Show that there is a bijection from $\mathbb{N}$ to $\set{\seq{n,m} : n,m \in \mathbb{N}}$, which is the set of pairs of natural numbers. (9~points)}

\answer{Use the same construction that was used to show that the natural numbers are in one-one correspondence with the rational numbers. In other words: arrange the pairs $\seq{n,m}$ in a table like that for the rationals, where $n$ increases rightward on the horizontal axis and $m$ increases downward on the vertical axis. Then count diagonally across the table just as we do to count the rationals. This gives us a bijection from the natural numbers to the set of pairs of natural numbers.}

}




%%% QUESTION LEFT OUT TO ADD VARIETY
\com{
\item \question{Is there a bijection between the natural numbers and the set of functions from natural numbers to natural numbers? (15 points)}

\answer{No. This is just a version of Cantor's proof that there is no bijection between the natural numbers and [0,1]. Ideally, the student would reproduce some version of that proof. }
}



\item \question{
Let $S$ be the following set: $$\set{f : f \text{ is a function that assigns a member of $\set{\Earth,\Sun,\Moon}$ to each natural number}}$$
where $\Earth$ is the Earth, $\Sun$ is the Sun, and $\Moon$ is the Moon.
An example of a member of $S$ is the function $g: \mathbb{N} \rightarrow \set{\Earth,\Sun,\Moon}$ such that $$g(n) = \begin{cases}\Earth \text{ if $n$ is a power of seven}\\ \Sun \text{ otherwise}\end{cases}$$
Show that there cannot be a bijection from the set of natural numbers to $S$. (9~points)


}

\answer{One can use a version of the construction that we used to show that there are more real numbers than natural numbers.

Assume for \emph{reductio} that there is a bijection $f$ from the natural numbers to $S$, and let $f(n)$ be the function $g_n: \mathbb{N} \rightarrow \set{\Earth,\Sun,\Moon}$. (Since $f$ is a bijection, this means that $S = \set{g_n : n \in \mathbb{N}}$.)


Now consider the `diagonal' sequence $<g_0(0),g_1(1),\ldots>$, and construct its evil twin $<t(g_0(0)),t(g_1(1)),\ldots>$, where $t(x) = \Sun$ if $x \neq \Sun$, and $t(x) = \Earth$ if $x = \Earth$. The evil twin function $h(n)=t(g_n(n))$ is a function from natural numbers to $S$. But $h \neq g_m$ for every $m$, since $g_m(m) \neq t(g_m(m)) = h(m)$. This contradicts our earlier claim that  $S = \set{g_n : n \in \mathbb{N}}$.}


\item
\question{The unit cube $C$ is the set of triples $\seq{r,p,q}$, for $r,p,q \in [0,1]$. A point $\seq{r,p,q}$ in $C$ is said to be ``rational'' if each of $r,p,$ and $q$ is rational.}

\begin{enumerate}

% For next time: people find part (a) confusing (they assume that C consists of only rational points), and the question is pretty close to something explicitly used in lecture. Best to replace (or eliminate) part (a) for next time.

\item \question{Is there a bijection from the set of points in $C$ to the set of real numbers? (8~points; don't forget to justify your answer)}

\answer{Yes. By exercise 3 of section 1.6.3, there is a bijection from $\mathbb{R}$ to $[0,1]$ and by exercise 6 of section 1.6.3, there is a bijection from $[0,1]$ to $C$. The result follows from transitivity.}


\item \question{Is there a bijection from the set of rational points in $C$ to the set of rational numbers? (8~points; don't forget to justify your answer)}

\answer{Yes. 

From a problem above, we know that there is a bijection $f$ from $\mathbb{N}\times\mathbb{N}$ to  $\mathbb{N}$. We can then define a bijection $g$ from $\mathbb{N}\times\mathbb{N} \times \mathbb{N}$ to  $\mathbb{N}$ as follows:
$$g(\seq{n,m,l}) = f(f(n,m),l)$$
Suitable variations of the diagonal construction can then be used to show that there is a bijection $h$ from $[0,1] \cap \mathbb{Q}$ to $\mathbb{N}$, and a bijection $i$ from $\mathbb{N}$ to $\mathbb{Q}$.

We can then define a conjunction $j$ from the set of rational points in $C$ to $\mathbb{Q}$. For $\seq{r,p,q}$ a rational point in $C$, define $j$ as follows:
$$
j(\seq{r,p,q}) = i(g(\seq{h(r),h(p),h(q)}))
$$




}




\end{enumerate}



\item \question{
Consider the following infinite tree:

\Tree [.  [.0 [.0 [.0 0 1 ] [.1 0 1 ] ] [.1 [.0 0 1 ] [.1 0 1 ] ] ] [.1 [.0 [.0 0 1 ] [.1 0 1 ] ] [.1 [.0 0 1 ] [.1 0 1 ] ] ] ]

\begin{center} \ \vdots \end{center}

(When fully spelled out, the tree contains one row for each natural number. The zero-th row contains one node, the first row contains two nodes, the second row contains four nodes, and, in general, the $n$th row contains $2^n$ nodes.)}

\begin{enumerate}

\item \question {Is there a bijection between the \emph{nodes} of this tree and the natural numbers? (8~points)}

\answer{There are as many nodes as natural numbers. One way to see this is to note that each node can be arrived at as a result of taking a finite number of 0-1 decisions as one travels down on the tree. Accordingly, each node corresponds to a unique finite sequence of zeroes and ones. 

So, we need to verify that there is a bijection between the set $B = \set{s : \text{$s$ is a finite binary sequence}}$ and $\mathbb{N}$. But one can generate such a bijection by enumerating the members of $B$:
$$
\seq{}, \seq{0}, \seq{1}, \seq{0,0}, \seq{0,1}, \seq{1,0}, \seq{1,1}, \seq{0,0,0},\dots
$$
and assigning the number $n$ to the $n$the member of the enumeration. (Think of the enumeration as having a zero-th member.)
}

\item \question{Is there a bijection between the \emph{paths} of this tree and the natural numbers? A path is an infinite sequence of nodes which starts at the top of the tree and contains a node at every row, with each node connected to its successor by an edge. (Paths can be represented as infinite sequences of zeroes and ones.)   (9~points; don't forget to justify your answers!)}

\answer{There are more paths than natural numbers. One way to see this is by noting that each path can be arrived at as a the result of taking infinitely many 0-1 decisions as one travels down the tree. So each node corresponds to an infinite sequence of zeroes and ones, or, more precisely, to a function from the natural numbers to the set $\{0,1\}$. But there is a bijection between the set of such functions and $\powerset(\mathbb{N})$, which we showed in the lecture notes to have the same cardinality as the real numbers.
}
\end{enumerate}

\item \question{Let $\mathscr{P}^\star(A)$ be the set of non-empty subsets of $A$.}

\begin{enumerate}

\item \question{Show that the following analogue of Cantor's Theorem is false: for any set $A$, $|A| < |\mathscr{P}^\star(A)|$. (2~points)}

\answer{$\set{0}$ has exactly one member and exactly one non-empty subset.}

\item \question{Show that the following analogue of Cantor's Theorem is true: for any set $A$ with two or more members, $|A| < |\mathscr{P}^\star(A)|$. (8~points)}

\answer{Suppose otherwise. Then there is some set $A$ such that $|A| \not< |\mathscr{P}^\star(A)|$, which means that either there is no injection from $A$ to $\mathscr{P}^\star(A)$ or there is a bijection from $A$  to $\mathscr{P}^\star(A)$. But $f(x) = \set{x}$ an injection from $A$ to $\mathscr{P}^\star(A)$. So it must be the case that there is a bijection $h'$ from $A$ to $\mathscr{P}^\star(A)$, and therefore a bijection $h$ from $\mathscr{P}^\star(A)$ to $A$.

Let $D = \set{a \in A : \exists C \in \mathscr{P}^\star(A) (h(C) = a \wedge a \not\in C)}$. We start by verifying that $D$ is non-empty. Suppose otherwise. Then $D$ is empty, so, in particular, $h(\set{a}) = a$ for each $a \in A$. But $A$ has at least two members, $a$ and $b$ ($a \neq b$). And, since $h$ is injective, the fact that $h(\set{a}) = a$ and $h(\set{b}) = b$ entails that we can have neither $ h(\set{a,b}) = a $ nor $h(\set{a,b}) = b$. So $h(\set{a,b}) \in D$, contradicting our hypothesis.

Now consider the question whether $h(D) \in D$ is true or false. If it's true, we have $h(C) = h(D) \wedge h(D) \not\in C$ for some $C \in \mathscr{}P^\star(A)$. Since $h$ is an injection, $h(C) = h(D)$ entails $C = D$, so we must have $h(D) \not\in D$, contradicting our hypothesis. If, on the other hand, $h(D) \in D$ is false, $h(C) = h(D) \wedge h(D) \not\in C$ must be false for every $C \in \mathscr{}P^\star(A)$. Since $D$ is non-empty, it follows that $h(C) = h(C) \wedge h(C) \not\in C$ must be false, which entails $h(C) \in C$, contradicting our hypothesis.
}

\end{enumerate}
\question{\emph{Notation:} $|A|$ is the cardinality of $A$ and $|A| < |B|$ means that there is an injection from $A$ to $B$ but no bijection from $A$ to $B$.}

\end{enumerate}



%Endnotes
%\newpage \begingroup \parindent 0pt \parskip 2ex \def\enotesize{\normalsize} \theendnotes \endgroup


%\newpage\bibliographystyle{linquiry}\bibliography{agustin}

%\end{doublespace}
\end{document}

